Long-form recording of infant body position in the home using wearable inertial sensors

### 2023-10-17 manuscript.Rmd
Hailey and Max edits
lab meeting notes and examples

---
title             : "Long-form recording of infant body position in the home using wearable inertial sensors"
shorttitle        : "Infant position in the home"

author: 
  - name          : "John M. Franchak"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "UC Riverside Department of Psychology, 900 University Avenue, Riverside, CA 92521"
    email         : "franchak@ucr.edu"
  - name          : "Maximilian Tang"
    affiliation   : "1"
  - name          : "Hailey Rousey"
    affiliation   : "1"
  - name          : "Chuan Luo"
    affiliation   : "1"

authornote: |
  We are grateful to Vanessa Scott, Tasnia Haider, and Ishapreet Kaur for their help in collecting data for the present study and to the research assistants of the UCR Perception, Action, and Development Lab for annotating videos.

  This work was funded by National Science Foundation Grant BCS #1941449 to the first author. The authors declare no conflict of interest.

abstract: |
  Long-form audio recordings have had a transformational effect on the study of infant language acquisition by using mobile, unobtrusive devices to gather full-day, real-time data that can be automatically scored. How can we produce similar data in service of measuring infants' everyday motor behaviors, such as body position? The aim of the current study was to validate long-form recordings of infant position (supine, prone, sitting, upright, held by caregiver) based on machine learning classification of data from inertial sensors worn on infants' ankles and thighs. Using over 100 hours of video recordings synchronized with inertial sensor data from infants in their homes, we demonstrate that body position classifications are sufficiently accurate to measure infant behavior. Moreover, classification remained accurate when predicting behavior later in the session when infants and caregivers were unsupervised and went about their normal activities, showing that the method can handle the challenge of measuring unconstrained, natural activity. Next, we show that the inertial sensing method has convergent validity by replicating age differences in body position found using other methods with full-day data captured from inertial sensors. We end the paper with a discussion of the novel opportunities that long-form motor recordings afford for understanding infant learning and development.

keywords          : "body position, motor development, everyday experiences, sitting, machine learning"
wordcount         : "11,995"

bibliography      : "master.bib"

floatsintext      : no
linenumbers       : no
draft             : no
mask              : no
figurelist        : no
tablelist         : no
footnotelist      : no
biblio-style      : "apa"
classoption       : "man"
output:
  papaja::apa6_pdf:
    citation_package: biblatex
header-includes:
  - \raggedbottom
editor_options: 
  chunk_output_type: console
---

@@

The first set of results showed that group and individual models trained from data during the proximal comparison period were accurate during the proximal period. In the next analysis, we examine long-term performance by testing how accurately models trained from the proximal period could predict body position during the distal comparison period. A single group model was created using all sessions' proximal period training data (rather than group models leaving out a single session); the same individual models were used. Distal videos had only coarse temporal synchrony with motion recordings which precluded calculating accuracy based on the proportion of matching events. Instead, we summed the amount of time infants were predicted to be in each of the 5 body position categories from the model and compared that to the summed time for the body positions based on human coding [@YaoPlotz2019; @FranchakScott2021].

>>>>

---
title             : "Long-form recording of infant body position in the home using wearable inertial sensors"
shorttitle        : "Infant position in the home"
author: 
  - name          : "John M. Franchak"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "UC Riverside Department of Psychology, 900 University Avenue, Riverside, CA 92521"
    email         : "franchak@ucr.edu"
  - name          : "Maximilian Tang"
    affiliation   : "1"
  - name          : "Hailey Rousey"
    affiliation   : "1"
  - name          : "Chuan Luo"
    affiliation   : "1"

authornote: |
  We are grateful to Vanessa Scott, Tasnia Haider, and Ishapreet Kaur for their help in collecting data for the present study and to the research assistants of the UCR Perception, Action, and Development Lab for annotating videos.
  This work was funded by National Science Foundation Grant BCS #1941449 to the first author. The authors declare no conflict of interest.
abstract: |
  Long-form audio recordings have had a transformational effect on the study of infant language acquisition by using mobile, unobtrusive devices to gather full-day, real-time data that can be automatically scored. How can we produce similar data in service of measuring infants' everyday motor behaviors, such as body position? The aim of the current study was to validate long-form recordings of infant position (supine, prone, sitting, upright, held by caregiver) based on machine learning classification of data from inertial sensors worn on infants' ankles and thighs. Using over 100 hours of video recordings synchronized with inertial sensor data from infants in their homes, we demonstrate that body position classifications are sufficiently accurate to measure infant behavior. Moreover, classification remained accurate when predicting behavior later in the session when infants and caregivers were unsupervised and went about their normal activities, showing that the method can handle the challenge of measuring unconstrained, natural activity. Next, we show that the inertial sensing method has convergent validity by replicating age differences in body position found using other methods with full-day data captured from inertial sensors. We end the paper with a discussion of the novel opportunities that long-form motor recordings afford for understanding infant learning and development.

keywords          : "body position, motor development, everyday experiences, sitting, machine learning"
wordcount         : "11,995"
bibliography      : "master.bib"
floatsintext      : no
linenumbers       : no
draft             : no
mask              : no
figurelist        : no
tablelist         : no
footnotelist      : no
biblio-style      : "apa"
classoption       : "man"
output:
  papaja::apa6_pdf:
    citation_package: biblatex
header-includes:
  - \raggedbottom
editor_options: 
  chunk_output_type: console
---

@@

The first set of results showed that group and individual models trained from data during the proximal comparison period were accurate during the proximal period. This suggests that the immediate accuracy of body position predictions early in the recording session was strong, but does not address how accurately predictions will be later in the recording. In the next analysis, we examine long-term performance by testing how accurately models trained from the proximal period could predict body position during the distal comparison period. A single group model was created using all sessions' proximal period training data (rather than group models leaving out a single session); the same individual models were used. Distal videos had only coarse temporal synchrony with motion recordings which precluded calculating accuracy based on the proportion of matching events. Instead, we summed the amount of time infants were predicted to be in each of the 5 body position categories from the model and compared that to the summed time for the body positions based on human coding [@YaoPlotz2019; @FranchakScott2021].
